---
title: "03 Combination"
author: "J Andres Gannon, Danil Gromakov, Katelyn Hess, and Yulia Krylova"
date: "May 2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(ggplot2)
```

This document produces some descriptive summary statistics of the new data collected regarding US prosecutor and Department of Justice officials appointed during the Trump administration

# Load data
```{r}
prosecutors <- read.csv(file = paste0(here::here(), '/data/prosecutors_final.csv'))
justicedept <- read.csv(file = paste0(here::here(), '/data/justicedept_final.csv'))
```

Combine into a single df
```{r}
# Check column names and fix for merging
sort(colnames(prosecutors))
sort(colnames(justicedept))

prosecutors <- prosecutors %>% dplyr::rename(american_center_for_law_and_justice = american_center_for_law_and_justice_tie,
                                             christian_legal_society_bar_association = christian_legal_society_bar_association_tie,
                                             position_current = district,
                                             employer_past = employer,
                                             federalist_society = federalist_society_tie,
                                             heritage_foundation = heritage_foundation_tie) %>%
  dplyr::select(-c("X"))
prosecutors$source <- "prosecutor"

justicedept <- justicedept %>% dplyr::rename(position_current = agency_title,
                                             employer_past = employer) %>%
                dplyr::select(-c("X", "title", "agency_name"))
justicedept$source <- "justicedept"

df <- dplyr::bind_rows(prosecutors, justicedept) %>%
  dplyr::select("person", "position_current", "source",
                "date_of_appointment",
                "date_of_birth", "age", "birthplace", "sex",
                "bachelors", "masters", "doctorate", "law", "law_conservative_score", "law_rank",
                "barr_appointment", "political_affiliation", 
                "employer_past", "firm_conservative_score", "start_date", "end_date",
                "american_center_for_law_and_justice", "christian_legal_society_bar_association", "federalist_society", "heritage_foundation", "national_rifle_association_tie", "republican_national_lawyers_association_tie")

df_noemploy <- df %>% dplyr::select(-c("employer_past", "firm_conservative_score", "start_date", "end_date")) %>%
  dplyr::distinct()
```

# Aggregate
Here, we use all the covariates gathered above to identify whether each person has a connection to Trump, Trump's allies, a conservative agenda, a career professional, or other

## Load
This df has the final aggregated list of unique persons from both datasets we have created
```{r, eval}
core <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                                    sheet = "all")

core <- janitor::clean_names(core)
```

## Conservative agenda
We code whether someone has a conservative agenda based on whether they attended a conservative law school, worked at a conservative law firm, have tied to conservative organizations, or are a conservative/Republican elected politician (state senator or state representative)

### Law school
```{r}
# Calculate average law school conservatism
school <- read.csv(file = paste0(here::here(), '/data/ideology_lawschool.csv'))
school$score_scaled <- scale(school$conservatism)

# Merge scaled score into a temp df
school_temp <- df_noemploy %>% dplyr::select(person, law, law_conservative_score) %>%
  dplyr::rename(school = law) %>%
  dplyr::distinct()

school_temp <- dplyr::left_join(school_temp, school)

# Create new column for worked at a more conservative than average firm
school_temp <- school_temp %>% dplyr::mutate(conservative = dplyr::if_else(score_scaled > 0, 1, 0)) %>%
  dplyr::select(person, conservative) %>%
  dplyr::rename(connection_conservative_lawschool = conservative) %>%
  dplyr::distinct()

# Merge that into core
tomerge <- school_temp
```

### Law firm
```{r}
# Calculate average law firm conservatism
firm <- read.csv(file = paste0(here::here(), '/data/ideology_lawfirm.csv'))
firm$score_scaled <- scale(firm$conservatism)

# Merge scaled score into a temp df
firm_temp <- df %>% dplyr::select(person, employer_past, firm_conservative_score) %>%
  dplyr::rename(firm = employer_past) %>%
  dplyr::distinct()

# Fix names for string matching
firm_temp$firm <- gsub(" LLP", "", firm_temp$firm)

firm_temp <- dplyr::left_join(firm_temp, firm)

# Create new column for worked at a more conservative than average firm
firm_temp <- firm_temp %>% dplyr::mutate(conservative = dplyr::if_else(score_scaled > 0, 1, 0)) %>%
  dplyr::select(person, conservative) %>%
  dplyr::rename(connection_conservative_firm = conservative)

firm_temp$connection_conservative_firm[is.na(firm_temp$connection_conservative_firm)] <- 0

firm_temp <- firm_temp %>% dplyr::distinct()

firm_temp <- firm_temp %>% dplyr::group_by(person) %>%
  dplyr::summarize(connection_conservative_firm = sum(connection_conservative_firm)) %>%
  dplyr::distinct()

# Merge that into core
tomerge <- dplyr::left_join(tomerge, firm_temp)
```

### Conservative membership
```{r}
# Calculate who has a 1 for either
memb_temp <- df_noemploy %>% dplyr::select(person, american_center_for_law_and_justice, 
                                           christian_legal_society_bar_association, 
                                           federalist_society, 
                                           heritage_foundation, 
                                           national_rifle_association_tie, 
                                           republican_national_lawyers_association_tie) %>%
                            dplyr::mutate(connection_conservative_group = ifelse(american_center_for_law_and_justice == 1 | 
                                            christian_legal_society_bar_association == 1 |
                                            federalist_society == 1 |
                                            heritage_foundation == 1 |
                                            national_rifle_association_tie == 1 |
                                            republican_national_lawyers_association_tie == 1,
                                            "1", "")) %>%
                           dplyr::select(person, connection_conservative_group) %>%
                           dplyr::distinct()

memb_temp$connection_conservative_group <- as.numeric(memb_temp$connection_conservative_group)

tomerge <- dplyr::left_join(tomerge, memb_temp)
```

### State senator/representative
```{r}
employ_temp <- df %>% dplyr::select(person, employer_past) %>%
  dplyr::distinct() %>%
  dplyr::mutate(connection_conservative_gov = ifelse(employer_past == "Idaho State Senate" |
                                          employer_past == "Tennessee State Senate" |
                                          employer_past == "Georgia House of Representatives" |
                                          employer_past == "Oklahoma House of Representatives",
                                          "1", "")) %>%
  dplyr::select(person, connection_conservative_gov) %>%
  dplyr::distinct()

employ_temp$connection_conservative_gov <- as.numeric(employ_temp$connection_conservative_gov)

employ_temp <- employ_temp %>% dplyr::group_by(person) %>%
  dplyr::summarize(connection_conservative_gov = sum(connection_conservative_gov)) %>%
  dplyr::distinct()

tomerge <- dplyr::left_join(tomerge, employ_temp)
```

### Merge
```{r}
# Combine tomerge into a single column called conservative_agenda_connections
tomerge <- tomerge %>% dplyr::mutate(connection_conservative_all = ifelse(connection_conservative_lawschool == 1 |
                                                                                connection_conservative_firm == 1 |
                                                                                connection_conservative_group == 1 |
                                                                                connection_conservative_gov == 1,
                                                                              "1", "0")) %>%
                                                                    dplyr::distinct()

tomerge$connection_conservative_all <- as.numeric(tomerge$connection_conservative_all)
tomerge$connection_conservative_all[is.na(tomerge$connection_conservative_all)] <- 0

# Merge into master list
final <- dplyr::left_join(core, tomerge)

# Replace NAs with 0
final[c("connection_trump_all", "connection_trump_interviews", "connection_trump_campaign", "connection_trumpally_all", "connection_trumpally_innercircle", "connection_trumpally_us_senate", "connection_trumpally_state_gov", "connection_trumpally_rnc", "connection_trumpally_administration", "connection_conservative_all", "connection_conservative_lawschool", "connection_conservative_firm", "connection_conservative_group", "connection_conservative_gov", "connection_careerprofessional_all", "connection_other_all")][is.na(final[c("connection_trump_all", "connection_trump_interviews", "connection_trump_campaign", "connection_trumpally_all", "connection_trumpally_innercircle", "connection_trumpally_rnc", "connection_trumpally_us_senate", "connection_trumpally_state_gov", "connection_trumpally_administration", "connection_conservative_all", "connection_conservative_lawschool", "connection_conservative_firm", "connection_conservative_group", "connection_conservative_gov", "connection_careerprofessional_all", "connection_other_all")])] <- 0

# Fix edge column NA conversions
final$connection_trumpally_us_senate[is.na(final$connection_trumpally_us_senate)] <- 1
final$connection_trumpally_state_gov[is.na(final$connection_trumpally_state_gov)] <- 1
final$connection_trumpally_rnc[is.na(final$connection_trumpally_rnc)] <- 1

# Merge everything else into that now
df_noemploy <- dplyr::left_join(final, df_noemploy)

# Reorder columns
df_noemploy <- df_noemploy %>% dplyr::select("person", "position_current", "source",
                "date_of_appointment",
                "date_of_birth", "age", "birthplace", "sex",
                "bachelors", "masters", "doctorate", "law", "law_conservative_score", "law_rank",
                "barr_appointment", "political_affiliation", 
                "connection_trump_all", "connection_trump_interviews", "connection_trump_campaign",
                "connection_trumpally_all", "connection_trumpally_innercircle", "connection_trumpally_us_senate", "connection_trumpally_state_gov", "connection_trumpally_rnc",
                "connection_conservative_all", "connection_conservative_lawschool", "connection_conservative_firm", "connection_conservative_group", "connection_conservative_gov",
                "connection_careerprofessional_all",
                "connection_other_all",
                "american_center_for_law_and_justice", "christian_legal_society_bar_association", "federalist_society", "heritage_foundation", "national_rifle_association_tie", "republican_national_lawyers_association_tie")
```

# Save
We have 2 versions of the data. One where the appointee is the unit of analysis and the other where prior employment for each appointtee is also listed, creating multiple rows per person.
```{r}
write.csv(df, paste0(here::here(), "/data/", "all_withemploy.csv"))
write.csv(df_noemploy, paste0(here::here(), "/data/", "all_noemploy.csv"))
```