---
title: "03 Combination"
author: "J Andres Gannon, Danil Gromakov, Katelyn Hess, and Yulia Krylova"
date: "May 2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(ggplot2)
```

This document produces some descriptive summary statistics of the new data collected regarding US prosecutor and Department of Justice officials appointed during the Trump administration

# Load data
```{r}
prosecutors <- read.csv(file = paste0(here::here(), '/data/prosecutors_final.csv'))
justicedept <- read.csv(file = paste0(here::here(), '/data/justicedept_final.csv'))
```

Combine into a single df
```{r}
# Check column names and fix for merging
sort(colnames(prosecutors))
sort(colnames(justicedept))

prosecutors <- prosecutors %>% dplyr::rename(american_center_for_law_and_justice = american_center_for_law_and_justice_tie,
                                             christian_legal_society_bar_association = christian_legal_society_bar_association_tie,
                                             position_current = district,
                                             employer_past = employer,
                                             federalist_society = federalist_society_tie,
                                             heritage_foundation = heritage_foundation_tie) %>%
  dplyr::select(-c("X"))
prosecutors$source <- "prosecutor"

justicedept <- justicedept %>% dplyr::rename(position_current = agency_title,
                                             employer_past = employer) %>%
  dplyr::select(-c("X", "title", "agency_name"))
justicedept$source <- "justicedept"

df <- dplyr::bind_rows(prosecutors, justicedept) %>%
  dplyr::select("person", "position_current", "source",
                "date_of_appointment",
                "date_of_birth", "age", "birthplace", "sex",
                "bachelors", "masters", "doctorate", "law", "law_conservative_score", "law_rank",
                "barr_appointment", "political_affiliation", 
                "employer_past", "start_date", "end_date",
                "american_center_for_law_and_justice", "christian_legal_society_bar_association", "federalist_society", "heritage_foundation", "national_rifle_association_tie", "republican_national_lawyers_association_tie")

df_noemploy <- df %>% dplyr::select(-c("employer_past", "start_date", "end_date")) %>%
  dplyr::distinct()
```

We have 2 versions of the data. One where the appointee is the unit of analysis and the other where prior employment for each appointtee is also listed, creating multiple rows per person.
```{r}
write.csv(df, paste0(here::here(), "/data/", "all_withemploy.csv"))
write.csv(df_noemploy, paste0(here::here(), "/data/", "all_noemploy.csv"))
```

# Aggregate
Here, we use all the covariates gathered above to identify whether each person has a connection to Trump, Trump's allies, a conservative agenda, a career professional, or other

## Load
This df has the final aggregated list of unique persons from both datasets we have created
```{r, eval}
core <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                                    sheet = "all")

core <- janitor::clean_names(core)
```

## Conservative agenda
We code whether someone has a conservative agenda based on whether they attended a conservative law school, worked at a conservative law firm, have tied to the federalist society or heritage foundation, or are a conservative/Republican elected politician (state senator or state representative)
```{r}

```

## Career professional
We code whether someone is a career professional based on whether they worked in the DOJ for at least 4 years

# Save
```{r, eval = FALSE}
# googlesheets4::write_sheet(all, ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
#                                    sheet = "all_noedit")
```

# Analyze data quality
## Data Explorer
```{r, fig.width = 16, fig.height = 16}
DataExplorer::plot_qq(df_noemploy)
```

## dlookr
```{r, fig.width = 12, fig.height = 16, warning=FALSE, error=FALSE}
dlookr::plot_normality(df_noemploy)
DT::datatable(dlookr::diagnose(df_noemploy))
```

## funModeling
```{r, fig.width = 16, fig.height = 16}
DT::datatable(funModeling::df_status(df_noemploy))
funModeling::plot_num(df_noemploy)
```

## naniar
```{r, fig.width = 16, fig.height = 8}
naniar::gg_miss_var(df_noemploy)
```

## smartEDA
```{r, fig.width = 16, fig.height = 16}
SmartEDA::ExpNumViz(df_noemploy)
```

## visdat
```{r, fig.width = 16, fig.height = 16}
visdat::vis_miss(df_noemploy, sort_miss = TRUE)
visdat::vis_dat(df_noemploy)
```