---
title: "Prosecutors"
author: "J Andres Gannon, Katelyn Hess, and Yulia Krylova"
date: "May 2020"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(ggplot2)
library(reticulate)
use_python("/home/andres/.local/share/r-miniconda/envs/r-reticulate/bin/python3.6.exe")
```

# Employment
## Googlesheet
### Load
We have 2 datasets for justice department employment info. The first is the google doc we have been manually curating and the second is the trump town employment data
```{r}
work <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                                    sheet = "employment_justicedepartment")
work <- janitor::clean_names(work)
```

### Clean
```{r}
# Check any notes
unique(work$notes)
work$notes <- NULL

# Subset to only senate appointments
work <- work %>% dplyr::filter(senate_appt == 1)

# Check employer strings
sort(unique(work$employer))

# Fix start and end date
work$start_date[work$start_date == "Unknown"] <- NA
work$end_date[work$end_date == "Unknown"] <- NA
```

## Trump town
### Load
```{r}
df_staff <- read.csv(file = paste0(here::here(), '/network_WH_outOrganisations/trumptown_data-store_1019/trumptown_staffer_organizations.csv'))
df_staff <- janitor::clean_names(df_staff)
```

### Clean
```{r}
# Subset to only attorneys
df_staff <- df_staff %>% dplyr::filter(stringr::str_detect(position_title_1, stringr::regex("attorney", ignore_case = TRUE)) | stringr::str_detect(position_title_1, stringr::regex("solicitor", ignore_case = TRUE)))

df_staff <- df_staff %>% dplyr::select(agency_name, position_title_1, staffer_name, name_organization, start_date, end_date, role_description) %>%
  dplyr::rename(position_title = position_title_1, attorney = staffer_name, employer = name_organization, title = role_description) %>%
  dplyr::distinct()

df_staff$start_date <- format(as.Date(df_staff$start_date, format = "%Y-%m-%d"), "%Y")
df_staff$start_date <- as.numeric(df_staff$start_date)

df_staff$end_date <- format(as.Date(df_staff$end_date, format = "%Y-%m-%d"), "%Y")
df_staff$end_date <- as.numeric(df_staff$end_date)

df_staff$notes <- NA
```

## Push to googlesheet to get list of senate appts
```{r}
df_staff <- df_staff %>% dplyr::select(agency_name, position_title, attorney) %>%
  dplyr::distinct()

df_staff$senateappt <- NA

googlesheets4::write_sheet(df_staff, ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                           sheet = "trumptown_names")
```

## Merge
```{r}
work_list <- as.data.frame(unique(work$attorney)) %>% dplyr::rename("attorney" = "unique(work$attorney)")
work_list$gsheet <- 1

staff_list <- as.data.frame(unique(df_staff$attorney)) %>% dplyr::rename("attorney" = "unique(df_staff$attorney)")

names <- dplyr::left_join(work_list, df_staff)
```

Combine the googlesheet justice department employment and the trump tower employment csv
```{r}
work_all <- dplyr::bind_rows(work, df_staff)

# Fix employer string mismatches
sort(unique(work_all$employer))
```

## Prep for network
```{r}
# Places of employment with the most unique attorneys
work_count <- work %>% dplyr::select(attorney, employer) %>%
  dplyr::distinct() %>%
  dplyr::group_by(employer) %>%
  dplyr::summarise(count = dplyr::n()) %>%
  dplyr::filter(count > 1)

DT::datatable(work_count)
```

## Save
````{r}
work <- work %>% dplyr::select(attorney, employer)
googlesheets4::write_sheet(work, ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                           sheet = "justicedepartmentprosecutors_noedit")

work <- write.csv(work, paste0(here::here(), "/network_WH_outOrganisations/", "justicedept_employment_hist.csv"))
```

# School
## Load
```{r}
edu <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                                    sheet = "school_justicedepartment",
                                 col_types = "ccccnnnnc")

edu <- janitor::clean_names(edu)
```

## Clean
```{r}
## Collapse degrees
sort(unique(edu$degree))

edu$degree[edu$degree == "AA" |
             edu$degree == "AB" |
             edu$degree == "BA" |
             edu$degree == "BBA" |
             edu$degree == "BS"] <- "bachelors"
 
edu$degree[edu$degree == "LLM" |
             edu$degree == "MA" |
             edu$degree == "MBA" |
             edu$degree == "MMgt" |
             edu$degree == "MPA" |
             edu$degree == "MPhil" |
             edu$degree == "MS"] <- "masters"

edu$degree[edu$degree == "PhD" |
             edu$degree == "EdD"] <- "doctorate"

edu$degree[edu$degree == "JD" |
             edu$degree == "LLB"] <- "law"

sort(unique(edu$degree))
```

## Prep for network
```{r}
# Make wide
edu_clean <- edu %>% dplyr::select(item_label, educated_at_label, degree) %>%
  tidyr::spread(degree, educated_at_label) %>%
  dplyr::select(item_label, bachelors, masters, law, doctorate)

edu_merge <- edu %>% dplyr::select(educated_at_label, law_rank, law_conservative_score)

edu_all <- dplyr::left_join(edu_clean, edu_merge, by = c("law" = "educated_at_label")) %>%
  dplyr::distinct()

# Clean each school col
sort(unique(edu_all$bachelors))
sort(unique(edu_all$masters))
sort(unique(edu_all$law))
```

## Results
```{r}
DT::datatable(edu_all)

GGally::ggpairs(edu_all[ , c('law_rank', 'law_conservative_score')], columnLabels = c("Law School Rank", "Law School Conservatism"))
ggplot2::ggsave("law_school_corr.png", path = paste0(here::here(), '/TrumpProsecutors_network'))
```

## Save
```{r}
googlesheets4::write_sheet(edu_all, ss = "https://docs.google.com/spreadsheets/d/1hPOBOyTujUReQzpjZ9PqIhj23V5C74EygMnmXIi6BgI/edit?usp=sharing",
                           sheet = "schooljusticedepartment_noedit")

edu <- write.csv(edu_all, paste0(here::here(), "/network_WH_outOrganisations/", "justicedept_edu_hist.csv"))
```
